A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-05-10 manuscript.Rmd
revamp simulation still WIP

---
title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
author: Aaron Peikert and Andreas M. Brandmaier 
output: bookdown::pdf_document2
repro:
  packages:
    - usethis
    - gert
    - aaronpeikert/repro@7bfaf98
    - here
    - rstudio/webshot2@f62e743
    - targets
    - renv
  scripts:
    - R/lavaan_helper.R
    - R/simulation.R
    - R/link.R
csl: apa7.csl
numbered_headings: true
bibliography: ["references.bib", "packages.bib"]
---

@@

```{r setup, include=FALSE}
library(repro)
library(tidyverse)
automate_load_scripts()
source(here::here("R", "link.R"))
```

@@

Multigroup Confirmatory Factor Analysis
First, establish strong measurement invariance using a CFI cutoff-criterion
Second, test for differences in latent factor means of Machiavellianism using a $\chi^2$ test

@@

```{r, include=FALSE}
source(here::here("R", "simulation.R"))
```

@@

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
The machine learning community has begun to use simulated data to test models and their implementation [@braiekTestingMachineLearning2020].
Beyond checking the technical side of the model, simulation is also beneficial on a conceptual leven [@axelrodAdvancingArtSimulation1997].
Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
If you identify unrealistic assumptions, you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical approach to address them.
We will show an example using structural equation modeling (SEM) for our Machiavellism example.
However, the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
One has to make many educated guesses to simulate data that closely matches your expectation.
These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
The analytic strategy (i.e. SEM) often gives you a hint about how to simulate data.

@@

```{r loadings, echo=FALSE}
knitr::kable(data.frame(Item = 1:9,
                        Loading = loadings_jones_paulhus),
             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
```

@@

@@

```{r, results='asis', echo=FALSE}
model_truth <- combine(
  measurement(
    "MACH",
    items("x", 9),
    loadings = round(loadings_jones_paulhus * 0.7, 2)
    ),
  intercepts("MACH", list(0, 0.2))
  )
cat("```r",
    model_truth,
    "```",
    sep = "\n")
model <- combine(measurement(
  "MACH",
  items("x", 9)),
  intercepts("MACH", list(0, "NA")))
```

@@

```{r, results='hide'}
simulated <- lavaan::simulateData(
  model_truth,
  sample.nobs = c(1000, 1000),
  meanstructure = TRUE,
  standardized = TRUE
)
```

@@

```{r}
fit <-
  lavaan::cfa(
    model,
    data = simulated,
    group = "group",
    meanstructure = TRUE,
    std.lv = TRUE
  ) 
```

Now, `lavaan` warns us that this model may not be identified.
We overlooked a crucial assumption of this model, that is, measurement invariance.
This assumption is also baked into our simulation of the data.
There is a good chance that this assumption is violated.
We, therefore, adapt our analysis to plan for this contingency.
After fitting the model with strict measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
For simplicity, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
Anyhow, remember that what we consider realistic for our case could be unfit for your case.

If you are satisfied with your assumptions, it helps to write a function that expresses them, like this one:
<!--Hands-on: build simple functions tailored for your analysis-->

@@

Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R", "GitHub")`.

With its help we can simulate 500 cases and fit our analysis.

```{r}
true_diff <- 0.2
model_truth <- combine(
  measurement(
    "MACH",
    items("x", 9),
    loadings = round(loadings_jones_paulhus * 0.7, 2)
  ),
  intercepts("MACH", list(0, true_diff)),
  variances("MACH", list(1, 1))
)

model_same <- 
  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
  MACH ~ c(0,0)*1
  MACH ~~ c(1, NA)*MACH"

model_differ <- 
  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
  MACH ~ c(0, NA)*1 + c(zero, diff)*1
  MACH ~~ c(1, NA)*MACH"
generate_data(500,
              model_truth,
              model_same,
              model_differ,
              list(std_estimate = get_std_estimate),
              group.equal = c("loadings", "intercepts"),
              mod_load = 3,
              mod_int = 3,
              std.lv = TRUE)
```

Note, that this certainly results in different numbers each time we run it.

```{r}
generate_data(500,
              model_truth,
              model_same,
              model_differ,
              list(std_estimate = get_std_estimate),
              group.equal = c("loadings", "intercepts"),
              mod_load = 3,
              mod_int = 3,
              std.lv = TRUE)
```

>>>>

---
title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
author: Aaron Peikert and Andreas M. Brandmaier 
output: bookdown::pdf_document2
repro:
  packages:
    - tidyverse
    - usethis
    - gert
    - aaronpeikert/repro@7bfaf98
    - here
    - rstudio/webshot2@f62e743
    - targets
    - renv
    - slider
    - patchwork
  scripts:
    - R/simulation.R
    - R/simulation_funs.R
    - R/link.R
csl: apa7.csl
numbered_headings: true
bibliography: ["references.bib", "packages.bib"]
---

@@

```{r setup, include=FALSE}
library(repro)
automate_load_packages()
automate_load_scripts()
source(here::here("R", "link.R"))
```

@@

Multigroup Confirmatory Factor Analysis with equal loadings and intercepts

@@

```{r, include=FALSE}
source(here::here("R", "simulation_funs.R"))
```

@@

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
The machine learning community has begun to use simulated data to test models and their implementation [@braiekTestingMachineLearning2020].
Beyond checking the technical side of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
If you identify unrealistic assumptions, you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical approach to address them.
We will show an example using structural equation modeling (SEM) for our Machiavellism example.
However, the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
One has to make many educated guesses to simulate data that closely matches your expectation.
These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
The analytic strategy (i.e. SEM) often gives you a hint about how to simulate data.

@@

```{r loadings, echo=FALSE}
knitr::kable(t(data.frame(Item = as.character(1:9),
                        Loading = loadings_jones_paulhus)),
             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
```

@@

```{r}
model_truth <-
"MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 + 0.43*x7 + 0.32*x8 + 0.36*x9
MACH ~ c(0, 0.2)*1
MACH ~~ c(1, 1)*MACH"
```

@@

@@

```{r}
simulated <- lavaan::simulateData(
  model_truth,
  sample.nobs = c(1000, 1000),
  meanstructure = TRUE,
  standardized = TRUE
)
```

@@

```{r}
model <- "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
MACH ~~ c(1, NA)*MACH
MACH ~ c(0, NA)*1 + c(zero, diff)*1"
fit <-
  lavaan::cfa(
    model,
    data = simulated,
    group = "group",
    meanstructure = TRUE,
    std.lv = TRUE
  ) 
```

Now, `lavaan` warns us that this model may not be identified.
We overlooked a crucial assumption of this model, that is, measurement invariance.
This assumption is also baked into our simulation of the data.
There is a good chance that this assumption is violated.
We, therefore, adapt our analysis to plan for this contingency.
After fitting the model with strong measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
For simplicity, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
Anyhow, remember that what we consider realistic for our case could be unfit for your case.

If you are satisfied with your assumptions for the data simulation, it helps to write a function that expresses them, like this one:
<!--Hands-on: build simple functions tailored for your analysis-->

@@

And to build another function that captures your planned analysis.
Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R", "GitHub")`.

```{r}
planed_analysis
extract_results
```

With its help we can simulate 1000 cases and fit our analysis.

```{r}
generate_data(1000) %>% 
  planed_analysis() %>% 
  extract_results()
```

Note, that this certainly results in different numbers each time we run it (see ["Randome Number Generation"](#prng)).
Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
First, you can check if the planned analysis recovers the simulated values (`std_estimate` should be close to the simulated mean difference of 0.2).
Second, it enables you to plan an appropriate sample size because you are able to quantify the sampling variability.

To complete a simulation study we need a function that repeats the analysis for different conditions (here only for different n):

```{r}
simulation_study
```

```{r}
res <- read_csv(here::here("results.csv"), col_types = "ddddd")
```

```{r}
res_interval <- res %>%
  select(n_obs, std_estimate) %>% 
  filter(!is.na(std_estimate)) %>% 
  group_by(n_obs) %>% 
  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
            .groups = "drop", median = median(std_estimate)) %>% 
  unnest(c(interval)) %>% 
  mutate(width = upper - lower, perc = 1 - alpha) %>% 
  group_by(alpha) %>% 
  arrange(n_obs) %>% 
  mutate(across(c(width, upper, lower),
                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
  ungroup() %>% 
  arrange(perc)
```

```{r}
choosen_perc <- .8
choosen_width <- .2
true_diff <- .2
minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
  filter(n_obs == min(n_obs))
sim_n <- pull(res, n_obs)
sim_nmax <- max(sim_n)
sim_n_unique <- unique(sim_n)
sim_nstep <- sim_n_unique[2] - sim_n_unique[1]
```

```{r}
plot_interval <- res_interval %>%
  ggplot(aes(
    n_obs,
    ymin = lower,
    ymax = upper,
    fill = perc,
    group = alpha
  )) +
  geom_ribbon() +
  geom_line(aes(y = median), color = "white") +
  geom_hline(yintercept = true_diff, linetype = "dotted") +
  scale_fill_viridis_c(
    begin = 1,
    end = 0,
    breaks  = 1 - seq(0.05, .2, 0.05),
    labels = . %>% scales::percent(accuracy = 1)
  ) +
  scale_y_continuous(n.breaks = 10) +
  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
  theme_minimal() +
  labs(x = "N per Group", y = "Standardized Effect Size") +
  theme(legend.title = element_blank()) +
  NULL
```

```{r}
plot_width <- res_interval %>%
  ggplot(aes(n_obs, width, color = perc, group = perc)) +
  geom_line() +
  geom_point(
    data = minn,
    size = 3,
    color = "black",
    shape = 3
  ) +
  scale_color_viridis_c(
    begin = 1,
    end = 0,
    breaks = 1 - seq(0.05, .2, 0.05),
    labels = . %>% scales::percent(accuracy = 1)
  ) +
  theme_minimal() +
  scale_y_log10(breaks = seq(0, 0.8, .1)) +
  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
  theme(legend.title = element_blank()) +
  NULL
```

```{r}
simulation_plot <- plot_interval +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
  plot_width +
  plot_layout(nrow = 2, guides = 'collect') +
  plot_annotation(tag_levels = 'A')
```

```{r}
simulation_plot
```

Plot A depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
The light yellow interval shows where 80% of the estimates land.
The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes.

Plot B represents the width of the interval shown in Plot A on a logarithmic scale.
The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.
